Global adjoint tomography -- Model GLAD-M25\footnote{This paper is dedicated to Dimitri Komatitsch, computational geophysicist extraordinaire, who passed away unexpectedly shortly before its submission.}

### 2019-02-07 main.tex
Updates from Overleaf

%\usepackage{subfig}

@@

\author[Lei et al.]
  {Wenjie Lei$^1$, Youyi Ruan$^{1.2}$, Ebru Bozda\u g$^3$, Daniel Peter$^4$, Matthieu Lefebvre$^1$, \\{\LARGE \rm Dimitri Komatitsch$^5$, Jeroen Tromp$^{1,6}$, Judith Hill$^7$, Norbert Podhorszki$^7$}, \\ {\LARGE \rm and David Pugmire$^7$} \\
  $^1$ Department of Geosciences, Princeton University, Princeton, NJ 08544, USA\\
  $^2$ Nanjing University, Nanjing, China\\
  $^3$ Department of Geophysics, Colorado School of Mines, Golden, CO 80401, USA\\
  $^4$ Extreme Computing Research Center, King Abdullah University of Science and Technology, \\Thuwal 23955-6900, Kingdom of Saudi Arabia\\
  $^5$ LMA, CNRS UPR 7051, Aix-Marseille University, Centrale Marseille, 13453 Marseille Cedex 13, France\\
  $^6$ Program in Applied and Computational Mathematics, Princeton University, Princeton, NJ 08544, USA\\
  $^7$ Oak Ridge National Laboratory, Oak Ridge, TN 37831, USA\\
  }

@@

%%---------------------------------------------
%
%   Summary
%
%%---------------------------------------------
\begin{summary}
Building on our first-generation global adjoint tomography model,
GLAD-M15, we present our second-generation transversely isotropic global model,
named GLAD-M25, which is the result of ten quasi-Newton tomographic iterations with an
earthquake database consisting of 1,480 events in the magnitude range $5.8\le M_w \le 7.0$,
an almost six-fold increase over the first-generation model.
We calculated fully 3D synthetic seismograms with a shortest period of 17~s based on a GPU-accelerated spectral-element wave
propagation solver which accommodates effects due to 3-D anelastic crust \& mantle structure, topography \& bathymetry, the ocean load, ellipticity, rotation, and self-gravitation.
We used an adjoint-state method to calculate Fr\'echet derivatives in 3D anelastic Earth models
facilitated by a parsimonious storage algorithm.
The simulations were performed on the Cray XK7 ``Titan'' at the Oak Ridge Leadership Computing Facility.
We quantitatively evaluated GLAD-M25 by assessing misfit reductions and traveltime \& amplitude anomaly histograms in twelve measurement
categories.
We performed similar assessments for a held-out data set consisting of 360 earthquakes,
with results comparable to the actual inversion.
We compare GLAD-M25 with numerous other global and regional models,
and highlight a variety of plumes and subduction zones.
\end{summary}

@@

This article is organized as follows.
Sections~\ref{section:start}--\ref{section:data} describe the starting model,
earthquake database, and seismographic data set. 
Sections~\ref{section:misfit} and~\ref{section:parameterization} describe the misfit function minimized during the inversion process
and the related model parameterization.
In Section~\ref{section:workflow} we describe the adjoint tomography workflow in some detail.
Section~\ref{section:misfit_evolution} describes the evolution of the misfit, and in Section~\ref{section:evaluation} we evaluate the model based on a held-out data set of 320 earthquakes.
Finally, in Section~\ref{section:model}, we describe the model in detail,
offering many comparisons with various global and regional models, as described in the proceeding paragraph.
We conclude with a discussion of future opportunities and directions.

@@

As a continuation of our previous work~\citep{bozdaug2016global},
we used first-generation model GLAD-M15 as our
starting model.
GLAD-M15 is a 3D transversely isotropic earth model, which combined
3D mantle model S362ANI~\citep{kustowski2008anisotropic}
with 3D crustal model CRUST2.0~\citep{bassin2000current} as its starting model.
In this study,
we continue to use the same transversely isotropic model parametrization.
Instead of relying on crustal corrections,
the mesh implementation of the crust in the spectral-element solver SPECFEM3D\_GLOBE~\citep{KoTr02a,KoTr02b,PeKoLuMaLeCaLeMaLiBlNiBaTr11} enables us to accurately accommodate topography and bathymetry as well as variations in the Moho.
GLAD-M15 was constructed using a global database of 256 earthquakes.
The first 12 iterations of the GLAD-M15 inversion used three-component seismograms with a shortest period of 27~s,
and the final 3 iterations reduced this further to 17~s.
In this study we also used three-component seismograms with a shortest period of 17~s.

@@

The seismographic stations used in this study were carefully selected to ensure global coverage
and high data quality (Fig.~\ref{fig:stations}).
In addition to the seismic stations used for the source inversions,
we included all available data for our 1,480 event earthquake database from many data centers,
including IRIS, ORFEUS, INGV, IPGP, ETH, and GEONET.
Regional and temporary networks,
such as US Array~(TA),
Africa Array~(AF), the Canadian National Seismograph Network~(CN), Geoscience Australia~(AU),
the Antarctic Seismographic Argentinean Italian Network~(AI),
and the New Zealand National Seismograph Network~(NZ),
constitute a significant part
of our database and greatly improve coverage in certain regions.

@@

With these considerations in mind,
the overall misfit, $\Phi$, is defined as follows:
\begin{equation}
\label{eq:misfit}
\Phi = \sum_{s}^{S} \omega_s \sum_{c}^{C} \omega_{c} \sum_{r}^{R_{sc}} \omega_{scr} \sum_{w}^{W_{scr}} \omega_{scrw}\, \chi_{scrw}
\quad ,
\end{equation}
where~$S$ denotes the number of sources, $C$ the number of categories,
$R_{sc}$ the number of receivers recording source~$s$ in category~$c$,
and~$W_{scr}$ the number of measurement windows for source~$s$, category~$c$,
and receiver~$r$.
The misfit for a specific source~$s$, category~$c$, receiver~$r$, and window~$w$ is
\begin{equation}
  \chi_{scrw} = \frac{1}{\Delta\omega}\int_{\omega_1}^{\omega_2} \Big( \frac {\Delta \tau_{scrw}} {\sigma_{scrw}} \Big)^2\, \mathrm{d}\omega
\quad ,
\end{equation}
where~$\Delta \tau_{scrw}$ denotes a multi-taper frequency-dependent phase measurement over the frequency interval~$\Delta\omega=\omega_2-\omega_1$\,,
with associated standard deviation~$\sigma_{scrw}$\,.
For body waves, the window misfit is simply the cross-correlation traveltime anomaly,
$\Delta T_{scrw}$\,, weighted by its standard deviation, i.e., $\chi_{scrw}=(\Delta T_{scrw}/\sigma_{scrw})^2$\,.
As in the source inversions,
the window weights are equal, $\omega_{scrw}=1$\,,
and the category weight, $\omega_c$\,, is the reciprocal of the number of measurements in that
category, $1/C$.
Earthquakes mainly occur along plate boundaries,
and most seismic stations are confined to the continents,
which leads to a very uneven distribution of sources and stations,
as illustrated in Figs.~\ref{fig:event_1480} and~\ref{fig:stations}.
Therefore, weighting is crucial in global tomography to balance this uneven sampling
of Earth's interior.
Our source and receiver weighting strategy,
encapsulated by the weights~$\omega_s$ and $\omega_{scr}$\,,
is developed to compensate for uneven spatial sampling.
Following the same strategy as for the receiver weights in the source inversions,
we define source and receiver weights as
\begin{equation}
\omega_{s}^{-1} = N\,\sum_{s'=1}^{S} \exp\left[\mbox{}-\left(\frac{\Delta_{ss'}}{\Delta}\right)^2\right]
\quad ,
\label{eq:source_weights}
\end{equation}
and
\begin{equation}
\omega_{scr}^{-1} = N_{sc}\,\sum_{r'=1}^{R_{sc}} \exp\left[\mbox{}-\left(\frac{\Delta_{rr'}}{\Delta_{sc}}\right)^2\right]
\quad ,
\label{eq:receiver_weights}
\end{equation}
respectively.
Here~$N$ and~$N_{sc}$ are normalization factors,
and~$\Delta_{ss'}$ and~$\Delta_{rr'}$ denote angular distances between source and receiver pairs~$\{s,s'\}$ and~$\{r,r'\}$\,.
The reference angular distances~$\Delta$ and~$\Delta_{sc}$ need to be chosen based on the condition 
numbers of the diagonal weighting matrices defined by Eqns.~(\ref{eq:source_weights}) and~(\ref{eq:receiver_weights}).

@@

As illustrated in Fig.~\ref{fig:preprocess_workflow}, the preprocessing workflow
includes the following phases.
\begin{enumerate}
  \item Signal processing to remove the instrument response from observed data
    to recover ground displacement. Both observed and synthetic data are bandpass
    filtered, and the horizontal components are rotated to obtain the radial and transverse components of motion.
  \item Window selection on a pair of
    observed and synthetic seismograms. Pyflex (a Python version
    of FLEXWIN) is
    used to automatically generate windows where observed and
    synthetic data are sufficiently close to make measurements based on user defined
    criteria.
  \item Cross-correlation traveltime or multi-taper phase measurements in selected windows.
  \item Window sorting and cleaning based on statistical analyses of all measurements to eliminate outliers.
  \item Preliminary adjoint source construction based on the sorted windows.
  \item Calculation and assignment of weights, as discussed in Section~\ref{section:misfit}.
  \item Construction of the final, properly weighted, adjoint sources.
\end{enumerate}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/misfit.pdf}
  \caption{\small{Evolution of the misfit function from GLAD-M15 to GLAD-M25.
  Each color denotes different stages of the inversion, as discussed in Sections~\ref{sec:I}--\ref{sec:III}.
  (a) Evolution of the total misfit. (b)--(j) Misfit in
  each period band (rows) on three components (columns). (e)--(g) At iteration 18,
  we  split the 40--100~s period band into two
  measurement categories: body waves (red diamonds) and surface waves
  (red squares).
  }}
  \label{fig:misfit}
\end{figure}

@@

%{\color{red} Wenjie: Make sure the variance reductions are calculated as ln (Phi\_M25 / Phi\_M15).}
%{\color{red} Wenjie: I was using (Phi\_M25 - Phi\_M15) / Phi\_M15. So let me modify it?}
%{\color{blue} Wenjie: yes please. It will only change things slightly, if at all. You need to do the same for Table~\ref{table:misfit_reduction_M15_M25_360}.}

@@

Another way to evaluate model performance is by assessing the distribution
of measurements in the various categories.
Again we use all 1,480 events and a set of identical windows on all three components
to assess GLAD-M25, GLAD-M15, and S362ANI combined with CRUST2.0.
The total number of selected windows exceeds 18~million.
Fig.~\ref{fig:phase_hist} shows histograms of the resulting phase
measurements in all twelve measurement categories.
We observe that the distributions generally become better centered on zero,
and that all standard deviations are steadily reduced.
Again we note that the inversion is aimed at reducing the overall misfit,
so there are small tradeoffs between different measurement categories.

@@

%{\color{red} Wenjie: Note: add spherical harmonic analysis later.}

@@

Asia, like Europe, has a complex tectonic history, involving
interactions among the Indian, Eurasian, Australian, Pacific, and Philippine
plates.
Map views of GLAD-M25 centered on Asia are plotted in Fig.~\ref{fig:asia-vs},
together with S362ANI$+$M, SL2013sv~\citep{SchaefferLebedev13}, and regional model EARA2014~\citep{chen2015multiparameter}.
EARA2014 is a transversely isotropic model based on adjoint
tomography, assimilating 1.7~million frequency-dependent traveltime
measurements from 227 earthquakes recorded by 1,869 seismic stations.
It has superior data coverage, mainly by incorporating data from the China Array.

@@

In this section we highlight various plume systems on Earth. Fig.~\ref{fig:plumes}
shows cross sections of several plumes, comparing models GLAD-M25 ($V_\textrm{P}$),
TX2015~\citep[$V_\textrm{S}$;][]{TX2015}, and SEMUCB-WM1~\citep[$V_\textrm{S}$;][]{french2015broad}.
The three models tend to agree near the CMB,
and there is overall good agreement between GLAD-M25 and TX2015,
but the models can differ substantially in the mid and upper mantle,
with profoundly different implications for mantle
convection.

In Fig.~\ref{fig:plumes}(a),
we see that the Afar plume in model GLAD-M25 originates from the CMB, narrows its diameter
in the mid mantle, and broadens again above the 660~km discontinuity.
It exhibits
very similar behavior in TX2015 but differs significantly in SEMUCB-WM1.

In Fig.~\ref{fig:plumes}(b), from left to right, we see in model GLAD-M25 1) the Farallon
slab; 2) the Bermuda hotspot above the 660~km discontinuity; 3) the Canary hotspot originating from the CMT and
rising all the way up the Earth's surface; 4) the Afar hotspot from another angle.

Fig.~\ref{fig:plumes}(c) features 1) South America subduction; 2) the Cape Verde hotspot extending from the CMB all the way to the surface; 3) the Hoggar hotspot from 660~km to the surface.

Fig.~\ref{fig:plumes}(d) shows a plume which rises from the CMB right below Iceland, reaches 660~km, amplifies,
and ascends to the surface.
The intensification above 660~km may be an indication of partial melting.

Fig.~\ref{fig:plumes}(e) contains 1) the Easter hotspot; 2) the Galapagos
hotspot; 3) the Central America slab.
In models GLAD-M25 and TX2015,
the Easter and Galapagos hotspots appear to originate from the same underlying plume in the deep mantle,
which is the eastern arm of the Pacific superplume.

Finally,
in Fig.~\ref{fig:plumes}(f) we see the Marion and Kerguelen hotspots, which might be connected to Afar in the lower mantle.
Recovering these plumes is challenging due to poor data coverage in the southern hemisphere.

@@

\begin{figure}
    \centering
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Aeg.png}\label{fig:subd_a}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Br.png}\label{fig:subd_b}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Ha.png}\label{fig:subd_c}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Ne.png}\label{fig:subd_d}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Wc.png}\label{fig:subd_e}}\\[-1pt]
    \caption{\small{Vertical cross sections of compressional wavespeed perturbations in various subduction zones
    for models GLAD-M25 ($V_\textrm{P}$; left column), GAP-P4~\citep{fukao2013subducted} (middle column) and UU-P07~\citep{van2018atlas} (right column).
    The map in the top left corner of each row shows the cross section with color-coded red, green, and white dots for geographical reference.
    The dashed black semicircles in the cross sections denote depths of 410~km, 660~km, and 1000~km.
    (a) Aegean; (b) South America; (c) Hatteras; (d) Nepal; (e)  Wichita.}}
    \label{fig:subd}
\end{figure}

In this section we highlight various subduction systems on Earth.
In Fig.~\ref{fig:subd} we compare GLAD-M25~($V_\textrm{P}$) with two compressional wavespeed models,
namely, GAP-P4~\citep{fukao2013subducted} and UU-P07~\citep{van2018atlas}.

Overall, the three models are in reasonably good agreement.
At shallow depths, model GLAD-M25 shows more distinct continental crust and cratons, thanks to the inclusion of surface waves.
In the upper and mid mantle,
the models share very similar features.
At depth greater than 1000km, 
we see a very strong expression of the ancient Farallon slab in model Glad-M25, which is much weaker in the two P models.
The difference may be attributed to increased data coverage thanks to USArray, as well as the use of finite-frequency Fr\'echet derivatives.

@@

Other future directions of research involve inversions for azimuthal anisotropy
and attenuation.
The feasibility of such inversions on a regional scale have already been demonstrated in a studies focused on Europe by~\cite{ZhuTromp2013} and \cite{Zhuetal2013}.

@@

In this appendix we determine the spherically symmetric (radial) average of model GLAD-M25.
To accomplish this,
we first need to transform the 3D spectral-element mesh --- which includes ellipticity, topography \& bathymetry, and undulations on internal boundaries --- into a spherical volume.
After these adjustments, we obtain a spectral-element mesh for a perfect cubed sphere~\citep{KoTr02a}.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/1d_profile.pdf}
  \caption{\small{1D radial shear and compressional wavespeed profiles for GLAD-M25, STW105, and PREM.
  The reference frequency for physical dispersion is~{\color{red} ??}~Hz.}}
  \label{fig:global-average}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth]{figures/1d_profile_ra.pdf}
  \caption{\small{1D radial-anisotropic shear and compressional wavespeed profiles for GLAD-M25 and STW105. The reference frequency for physical dispersion is~{\color{red} ??}~Hz.}}
  \label{fig:global-ra-average}
\end{figure}

@@

To accomplish the transformation,
we first need to transform the 3D mesh --- which includes ellipticity, topography \& bathymetry, and undulations on internal boundaries --- into a spherical volume.
After these adjustments, we obtain a spectral element mesh for a perfect sphere,
that is, the sort of mesh used for spherically symmetric earth models, such as PREM~\citep{PREM}.

>>>>

@@

\author[Lei et al.]
  {Wenjie Lei$^1$, Youyi Ruan$^{1,2}$, Ebru Bozda\u g$^3$, Daniel Peter$^4$, Matthieu Lefebvre$^1$, \\{\LARGE \rm Dimitri Komatitsch$^5$, Jeroen Tromp$^{1,6}$, Judith Hill$^7$, Norbert Podhorszki$^7$}, \\ {\LARGE \rm and David Pugmire$^7$} \\
  $^1$ Department of Geosciences, Princeton University, Princeton, NJ 08544, USA\\
  $^2$ School of Earth Sciences and Engineering, Nanjing University, Nanjing, Jiangsu 210023, China\\
  $^3$ Department of Geophysics, Colorado School of Mines, Golden, CO 80401, USA\\
  $^4$ Extreme Computing Research Center, King Abdullah University of Science and Technology (KAUST), \\Thuwal 23955-6900, Kingdom of Saudi Arabia\\
  $^5$ LMA, CNRS UPR 7051, Aix-Marseille University, Centrale Marseille, 13453 Marseille Cedex 13, France\\
  $^6$ Program in Applied and Computational Mathematics, Princeton University, Princeton, NJ 08544, USA\\
  $^7$ Oak Ridge National Laboratory, Oak Ridge, TN 37831, USA\\
  }

@@

%%---------------------------------------------
%
%   Summary
%
%%---------------------------------------------
\begin{summary}
Building on our first-generation global adjoint tomography model,
GLAD-M15, we present our second-generation transversely isotropic global model,
named GLAD-M25, which is the result of ten quasi-Newton tomographic iterations with an
earthquake database consisting of 1,480 events in the magnitude range $5.5\le M_w \le 7.2$,
an almost six-fold increase over the first-generation model.
We calculated fully 3D synthetic seismograms with a shortest period of 17~s based on a GPU-accelerated spectral-element wave
propagation solver which accommodates effects due to 3-D anelastic crust \& mantle structure, topography \& bathymetry, the ocean load, ellipticity, rotation, and self-gravitation.
We used an adjoint-state method to calculate Fr\'echet derivatives in 3D anelastic Earth models
facilitated by a parsimonious storage algorithm.
The simulations were performed on the Cray XK7 ``Titan'' at the Oak Ridge Leadership Computing Facility.
We quantitatively evaluated GLAD-M25 by assessing misfit reductions and traveltime \& amplitude anomaly histograms in twelve measurement
categories.
We performed similar assessments for a held-out data set consisting of 360 earthquakes,
with results comparable to the actual inversion.
We compare GLAD-M25 with numerous other global and regional models,
and highlight a variety of plumes and subduction zones.
\end{summary}

@@

This article is organized as follows.
Sections~\ref{section:start}--\ref{section:data} describe the starting model,
earthquake database, and seismographic data set. 
Sections~\ref{section:misfit} and~\ref{section:parameterization} describe the misfit function minimized during the inversion process
and the related model parameterization.
In Section~\ref{section:workflow} we describe the adjoint tomography workflow in some detail.
Section~\ref{section:misfit_evolution} describes the evolution of the misfit, and in Section~\ref{section:evaluation} we evaluate the model based on a held-out data set of 360 earthquakes.
Finally, in Section~\ref{section:model}, we describe the model in detail,
offering many comparisons with various global and regional models, as described in the proceeding paragraph.
We conclude with a discussion of future opportunities and directions.

@@

As a continuation of our previous work~\citep{bozdaug2016global},
we used the first-generation model GLAD-M15 as our
starting model.
GLAD-M15 is a 3D transversely isotropic earth model, which combined
3D mantle model S362ANI~\citep{kustowski2008anisotropic}
with 3D crustal model CRUST2.0~\citep{bassin2000current} as its starting model.
In this study,
we continue to use the same transversely isotropic model parametrization.
Instead of relying on crustal corrections,
the mesh implementation of the crust in the spectral-element solver SPECFEM3D\_GLOBE~\citep{KoTr02a,KoTr02b,PeKoLuMaLeCaLeMaLiBlNiBaTr11} enables us to accurately accommodate topography and bathymetry as well as variations in the Moho.
GLAD-M15 was constructed using a global database of 256 earthquakes.
The first 12 iterations of the GLAD-M15 inversion used three-component seismograms with a shortest period of 27~s,
and the final 3 iterations reduced this further to 17~s.
In this study we also used three-component seismograms with a shortest period of 17~s.

@@

The seismographic stations used in this study were carefully selected to ensure global coverage
and high data quality (Fig.~\ref{fig:stations}).
In addition to the seismographic  stations used for the source inversions,
we included all available data for our 1,480 event earthquake database from many data centers,
including IRIS, ORFEUS, INGV, IPGP, ETH, and GEONET.
Regional and temporary networks,
such as US Array~(TA),
Africa Array~(AF), the Canadian National Seismograph Network~(CN), Geoscience Australia~(AU),
the Antarctic Seismographic Argentinean Italian Network~(AI),
and the New Zealand National Seismograph Network~(NZ),
constitute a significant part
of our database and greatly improve coverage in certain regions.

@@

With these considerations in mind,
the overall misfit, $\Phi$, is defined as follows:
\begin{equation}
\label{eq:misfit}
\Phi = \sum_{s}^{S} \omega_s \sum_{c}^{C} \omega_{c} \sum_{r}^{R_{sc}} \omega_{scr} \sum_{w}^{W_{scr}} \omega_{scrw}\, \chi_{scrw}
\quad ,
\end{equation}
where~$S$ denotes the number of sources, $C$ the number of categories,
$R_{sc}$ the number of receivers recording source~$s$ in category~$c$,
and~$W_{scr}$ the number of measurement windows for source~$s$, category~$c$,
and receiver~$r$.
The misfit for a specific source~$s$, category~$c$, receiver~$r$, and window~$w$ is
\begin{equation}
  \chi_{scrw} = \frac{1}{\Delta\omega}\int_{\omega_1}^{\omega_2} \Big( \frac {\Delta \tau_{scrw}} {\sigma_{scrw}} \Big)^2\, \mathrm{d}\omega
\quad ,
\end{equation}
where~$\Delta \tau_{scrw}$ denotes a multi-taper frequency-dependent phase measurement over the frequency interval~$\Delta\omega=\omega_2-\omega_1$\,,
with associated standard deviation~$\sigma_{scrw}$\,.
For body waves, the window misfit is simply the cross-correlation traveltime anomaly,
$\Delta T_{scrw}$\,, weighted by its standard deviation, i.e., $\chi_{scrw}=(\Delta T_{scrw}/\sigma_{scrw})^2$\,.
As in the source inversions,
the window weights are equal, $\omega_{scrw}=1$\,,
and the category weight, $\omega_c$\,, is the reciprocal of the number of measurements in that
category, $1/C$.
Earthquakes mainly occur along plate boundaries,
and most seismographic  stations are confined to the continents,
which leads to a very uneven distribution of sources and stations,
as illustrated in Figs.~\ref{fig:event_1480} and~\ref{fig:stations}.
Therefore, weighting is crucial in global tomography to balance this uneven sampling
of Earth's interior.
Our source and receiver weighting strategy,
encapsulated by the weights~$\omega_s$ and $\omega_{scr}$\,,
is developed to compensate for uneven spatial sampling.
Following the same strategy as for the receiver weights in the source inversions,
we define source and receiver weights as
\begin{equation}
\omega_{s}^{-1} = N\,\sum_{s'=1}^{S} \exp\left[\mbox{}-\left(\frac{\Delta_{ss'}}{\Delta}\right)^2\right]
\quad ,
\label{eq:source_weights}
\end{equation}
and
\begin{equation}
\omega_{scr}^{-1} = N_{sc}\,\sum_{r'=1}^{R_{sc}} \exp\left[\mbox{}-\left(\frac{\Delta_{rr'}}{\Delta_{sc}}\right)^2\right]
\quad ,
\label{eq:receiver_weights}
\end{equation}
respectively.
Here~$N$ and~$N_{sc}$ are normalization factors,
and~$\Delta_{ss'}$ and~$\Delta_{rr'}$ denote angular distances between source and receiver pairs~$\{s,s'\}$ and~$\{r,r'\}$\,.
The reference angular distances~$\Delta$ and~$\Delta_{sc}$ need to be chosen based on the condition 
numbers of the diagonal weighting matrices defined by Eqns.~(\ref{eq:source_weights}) and~(\ref{eq:receiver_weights}).

@@

As illustrated in Fig.~\ref{fig:preprocess_workflow}, the preprocessing workflow
includes the following phases.
\begin{enumerate}
  \item Signal processing to remove the instrument response from observed data
    to recover ground displacement. Both observed and synthetic data are bandpass
    filtered, and the horizontal components are rotated to obtain the radial and transverse components of motion.
  \item Window selection on a pair of
    observed and synthetic seismograms. Pyflex (a Python version
    of FLEXWIN by \citet{krischer2015}) is
    used to automatically generate windows where observed and
    synthetic data are sufficiently close to make measurements based on user defined
    criteria.
  \item Cross-correlation traveltime or multi-taper phase measurements in selected windows.
  \item Window sorting and cleaning based on statistical analyses of all measurements to eliminate outliers.
  \item Preliminary adjoint source construction based on the sorted windows.
  \item Calculation and assignment of weights, as discussed in Section~\ref{section:misfit}.
  \item Construction of the final, properly weighted, adjoint sources.
\end{enumerate}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/misfit.pdf}
  \caption{\small{Evolution of the misfit function from GLAD-M15 to GLAD-M25.
  Each color denotes different stages of the inversion, as discussed in Sections~\ref{sec:I}--\ref{sec:III}.
  (a) Evolution of the total misfit. (b)--(j) Misfit in
  each period band (rows) on three components (columns). (e)--(g) Since iteration 18,
  we  split the 40--100~s period band into two
  measurement categories: body waves (diamonds) and surface waves
  (squares).
  }}
  \label{fig:misfit}
\end{figure}

@@

@@

Another way to evaluate model performance is by assessing the distribution
of measurements in the various categories.
Again we use all 1,480 events and a set of identical windows on all three components
to assess GLAD-M25, GLAD-M15, and S362ANI combined with CRUST2.0.
The total number of selected windows,
determined using model GLAD-M25, exceeds 18~million.
Fig.~\ref{fig:phase_hist} shows histograms of the resulting phase
measurements in all twelve measurement categories.
We observe that the distributions generally become better centered on zero,
and that all standard deviations are steadily reduced.
Again we note that the inversion is aimed at reducing the overall misfit,
so there are small tradeoffs between different measurement categories.

@@

@@

Asia, like Europe, has a complex tectonic history, involving
interactions among the Indian, Eurasian, Australian, Pacific, and Philippine
plates.
Map views of GLAD-M25 centered on Asia are plotted in Fig.~\ref{fig:asia-vs},
together with S362ANI$+$M, SL2013sv~\citep{SchaefferLebedev13}, and regional model EARA2014~\citep{chen2015multiparameter}.
EARA2014 is a transversely isotropic model based on adjoint
tomography, assimilating 1.7~million frequency-dependent traveltime
measurements from 227 earthquakes recorded by 1,869 seismographic  stations.
It has superior data coverage, mainly by incorporating data from the China Array.

@@

In this section we highlight various plume systems on Earth. Fig.~\ref{fig:plumes}
shows cross sections of several plumes, comparing models GLAD-M25 ($V_\textrm{P}$),
TX2015~\citep[$V_\textrm{S}$;][]{TX2015}, and SEMUCB-WM1~\citep[$V_\textrm{S}$;][]{french2015broad}.
The three models tend to agree near the CMB,
and there is overall good agreement between GLAD-M25 and TX2015,
but the models can differ substantially in the mid and upper mantle,
with profoundly different implications for mantle
convection:

\begin{itemize}
\item[]{\it Afar} In Fig.~\ref{fig:plumes}(a),
we see that the Afar plume in model GLAD-M25 originates from the CMB, narrows its diameter
in the mid mantle, and broadens again above the 660~km discontinuity.
It exhibits
very similar behavior in TX2015 but differs significantly in SEMUCB-WM1.  
\item[]{\it Bermuda and Canary} In Fig.~\ref{fig:plumes}(b), from left to right, we see in model GLAD-M25 (i) the Farallon
slab; (ii) the Bermuda hotspot above the 660~km discontinuity; (iii) the Canary hotspot originating from the CMT and
rising all the way up the Earth's surface; (iv) the Afar hotspot from another angle.
\item[]{\it Cape Verde and Hoggar} Fig.~\ref{fig:plumes}(c) features (i) South America subduction; (ii) the Cape Verde hotspot extending from the CMB all the way to the surface; (iii) the Hoggar hotspot from 660~km to the surface. 
\item[]{\it Iceland and Eifel} Fig.~\ref{fig:plumes}(d) shows a plume which rises from the CMB right below Iceland, reaches 660~km, amplifies,
and ascends to the surface.
The intensification above 660~km may be an indication of partial melting.
\item[]{\it Easter and Galapagos} Fig.~\ref{fig:plumes}(e) contains (i) the Easter hotspot; (ii) the Galapagos
hotspot; (iii) the Central America slab.
In models GLAD-M25 and TX2015,
the Easter and Galapagos hotspots appear to originate from the same underlying plume in the deep mantle,
which is the eastern arm of the Pacific superplume.
\item[]{\it Marion and Kerguelen}
Finally,
in Fig.~\ref{fig:plumes}(f) we see the Marion and Kerguelen hotspots, which might be connected to Afar in the lower mantle.
Recovering these plumes is challenging due to poor data coverage in the southern hemisphere.
\end{itemize}

@@

\begin{figure}
    \centering
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Aeg.png}\label{fig:subd_a}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Br.png}\label{fig:subd_b}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Ha.png}\label{fig:subd_c}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Wc.png}\label{fig:subd_e}}\\[-1pt]
    \sidesubfloat[]{\includegraphics[width=0.88\textwidth]{figures/subductions/Ne.png}\label{fig:subd_d}}\\[-1pt]
    \caption{\small{Vertical cross sections of compressional wavespeed perturbations in various subduction zones
    for models GLAD-M25 ($V_\textrm{P}$; left column), GAP-P4~\citep{fukao2013subducted} (middle column) and UU-P07~\citep{van2018atlas} (right column).
    The map in the top left corner of each row shows the cross section with color-coded red, green, and white dots for geographical reference.
    The dashed black semicircles in the cross sections denote depths of 410~km, 660~km, and 1000~km.
    (a) Aegean; (b) South America; (c) Hatteras; (d) Wichita. (e) Nepal; }}
    \label{fig:subd}
\end{figure}

In this section we highlight various subduction systems on Earth.
In Fig.~\ref{fig:subd} we compare GLAD-M25~($V_\textrm{P}$) with two compressional wavespeed models,
namely, GAP-P4~\citep{fukao2013subducted} and UU-P07~\citep{van2018atlas}.
Fig.~\ref{fig:subd}(a) features the Aegean subduction zone. All models show a slab penetrating into the lower mantle beyond a depth of 1,000~km.
Fig.~\ref{fig:subd}(b) shows the South American slab, which, in this cross section, ponds in the shallow lower mantle in all three models, before sinking all the way to the CMB.
Figs.~\ref{fig:subd}(c) and~(d) show cuts across the ancient Farallon slab, which penetrates deeply into the lower mantle.
At depths greater than 1,000~km, 
we see a very strong expression of the slab in model Glad-M25, which is much weaker in the two P models.
The difference may be attributed to increased data coverage thanks to USArray, as well as the use of finite-frequency Fr\'echet derivatives.
Finally, Fig.~\ref{fig:subd}(e) shows subduction beneath the Himalayas.

Overall, the three models are in reasonably good agreement.
At shallow depths, model GLAD-M25 shows more distinct continental crust and cratons, thanks to the inclusion of surface waves.
But in the upper and mid mantle,
the models share very similar attributes.

@@

Other future directions of research involve inversions for azimuthal anisotropy
and attenuation.
The feasibility of such inversions on a regional scale has already been demonstrated in studies focused on Europe by~\cite{ZhuTromp2013} and \cite{Zhuetal2013}.

@@

In this appendix we determine the spherically symmetric (radial) average of model GLAD-M25.
To accomplish this,
we first need to transform the 3D spectral-element mesh --- which includes ellipticity, topography/bathymetry and undulations on internal boundaries --- into a spherical volume.
After these adjustments, we obtain a spectral-element mesh for a perfect cubed sphere~\citep{KoTr02a}.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/1d_profile.pdf}
  \caption{\small{1D radial shear and compressional wavespeed profiles for GLAD-M25, STW105, and PREM.
  The reference frequency for physical dispersion is ~7.77~mHz.}}
  \label{fig:global-average}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth]{figures/1d_profile_ra.pdf}
  \caption{\small{1D radial-anisotropic shear and compressional wavespeed profiles for GLAD-M25 and STW105. The reference frequency for physical dispersion is ~7.77~mHz.}}
  \label{fig:global-ra-average}
\end{figure}

@@

To accomplish the transformation,
we first need to transform the 3D mesh --- which includes ellipticity, topography/bathymetry and undulations on internal boundaries --- into a spherical volume.
After these adjustments, we obtain a spectral element mesh for a perfect sphere,
that is, the sort of mesh used for spherically symmetric earth models, such as PREM~\citep{PREM}.
